Quantum Mechanics Assignment Two

### 2016-01-20 quantum/latex/a2.tex
b la

\begin{align*}
  <p^2> &= \int_{-\infty}^{\infty} \psi^* \left(-\hbar^2 \frac{d^2}{dx^2}\right)\psi dx\\
  &= \sqrt{\frac{m\omega}{\pi\hbar}}\int_{-\infty}^{\infty} -\hbar^2e^{\frac{-m\omega x^2}{\hbar}}
  \left(\frac{m^2\omega^2x^2}{\hbar^2} - \frac{m\omega x}{\hbar}\right)dx\\
  &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}\left(
  \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m^2\omega^2x^2}{\hbar^2}dx -
  \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m\omega}{\hbar}\right)dx\\
  &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}
  \left(\frac{1}{4}\sqrt{\frac{m\pi\omega}{\hbar}} -
  \frac{1}{2}\sqrt{\frac{m\pi\omega}{\hbar}}\right) \hspace{2cm}\mbox{Wolfram}\\
  &= \frac{\hbar m\omega}{4}\\
\end{align*}
This answer is intuitive, since increasing mass as well as oscillatory speed should both increase
the expected momentum.

@@

\begin{align*}
  <p> &= \bra{n}\hat{p}\ket{n}\\
  &= i\sqrt{\frac{m\hbar\omega}{2}}\bra{n}\left(a^\dagger-a\right)\ket{n}\\
  &= i\sqrt{\frac{m\hbar\omega}{2}}
  \left(\sqrt{n+1}\bra{n}\ket{n+1} - \sqrt{n}\bra{n}\ket{n-1}\right) = 0\\
\end{align*}
This is an interesting result. It is intuitive that any wave function should have an expectation of
zero for momentum, else the wave function would move through space.

How does this reconcile w/ fact that p involves a spacial derivative?

\end{document}

>>>>

\begin{align*}
  <p^2> &= \int_{-\infty}^{\infty} \psi^* \left(-\hbar^2 \frac{d^2}{dx^2}\right)\psi dx\\
  &= \sqrt{\frac{m\omega}{\pi\hbar}}\int_{-\infty}^{\infty} -\hbar^2e^{\frac{-m\omega x^2}{\hbar}}
  \left(\frac{m^2\omega^2x^2}{\hbar^2} - \frac{m\omega}{\hbar}\right)dx\\
  &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}\left(
  \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m^2\omega^2x^2}{\hbar^2}dx -
  \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m\omega}{\hbar}\right)dx\\
  &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}
  \left(\frac{1}{2}\sqrt{\frac{m\pi\omega}{\hbar}} -
  \sqrt{\frac{m\pi\omega}{\hbar}}\right) \hspace{2cm}\mbox{Wolfram}\\
  &= \frac{\hbar m\omega}{2}\\
\end{align*}
This answer is intuitive, since increasing mass as well as oscillatory speed should both increase
the expected momentum.

@@

\begin{align*}
  <p> &= \bra{n}\hat{p}\ket{n}\\
  &= i\sqrt{\frac{m\hbar\omega}{2}}\bra{n}\left(a^\dagger-a\right)\ket{n}\\
  &= i\sqrt{\frac{m\hbar\omega}{2}}
  \left(\sqrt{n+1}\bra{n}\ket{n+1} - \sqrt{n}\bra{n}\ket{n-1}\right) = 0\\
\end{align*}
This is an interesting result. It is intuitive that any wave function should have an expectation of
zero for momentum, else the wave function would move through space due to a nonzero average
velocity.

\begin{align*}
  <x^2> &= \bra{n}x^2\ket{n}\\
  &= \frac{\hbar}{2m\omega}\bra{n}\left(a + a^\dagger\right)\left(a + a^\dagger\right)\ket{n}\\
  &= \frac{\hbar}{2m\omega}\bra{n}
  \left(aa + aa^\dagger + a^\dagger a + a^\dagger a^\dagger\right)\ket{n}\\
  &= \frac{\hbar}{2m\omega}
  \left(aa + aa^\dagger + a^\dagger a + a^\dagger a^\dagger\right)\ket{n}\\
  &= \frac{\hbar}{2m\omega} \left(\sqrt{n-1}\sqrt{n-2}\bra{n}\ket{n-2} + (n+1)\bra{n}\ket{n} +
  (n)\bra{n}\ket{n} + \sqrt{n+1}\sqrt{n+2}\bra{n}\ket{n+2}\right)\\
  &= \frac{\hbar}{2m\omega} \left(2n+1\right)\\
\end{align*}
This is an interesting result which indicates that for each extra rung on the ladder, the
expectation of the squared position increases linearly, meaning the expectation of the magnitude
of x increases as $\sqrt{n}$. This makes sense, since increasing energy levels have less and less
room to expand in the quadratic potential well. The $n=0$ state corresponds to the wave function
solution for the ground state found above.\\

\begin{align*}
  <p^2> &= \bra{n}p^2\ket{n}\\
  &= -\frac{m\hbar\omega}{2}\bra{n}\left(a^\dagger-a\right)\left(a^\dagger-a\right)\ket{n}\\
  &= -\frac{m\hbar\omega}{2}\bra{n}
  \left(a^\dagger a^\dagger - a^\dagger a - aa^\dagger + aa\right)\ket{n}\\
  &= -\frac{m\hbar\omega}{2}\bra{n}\left(\sqrt{n-1}\sqrt{n-2}\bra{n}\ket{n-2}
  - (n)\bra{n}\ket{n} - (n+1)\bra{n}\ket{n} + \sqrt{n+1}\sqrt{n+2}\bra{n}\ket{n+2}\right)\\
  &= \frac{\hbar m\omega}{2}\left(2n + 1\right)
\end{align*}
This result is similar to the above one for $x^2$ with different units. It is still a $\sqrt{n}$
relationship for momentum, which makes sense.\\

\textbf{c.) Uncertainty principle}\\
Since the above two cases agree with each other, we check that the more general case of part
\textbf{b.} agrees with the uncertainty principle.

\begin{align*}
  \Delta x &= \sqrt{<x^2>-<x>^2}\\
  &= \sqrt{\frac{\hbar}{2m\omega} \left(2n+1\right)}\\
  \Delta p &= \sqrt{<p^2>-<p>^2}\\
  &= \sqrt{\frac{\hbar m\omega}{2}\left(2n + 1\right)}\\
  \Delta x \Delta p &= \sqrt{\frac{\hbar}{2m\omega} \left(2n+1\right)}
  \sqrt{\frac{\hbar m\omega}{2}\left(2n + 1\right)}\\
  &= \frac{\hbar}{2}(2n+1)\\
\end{align*}
This is an interesting result which indicates that uncertainty increases as the pure wave function
gets more energetic. Ground state is the state with maximum certainty. The higher states get
progressively more uncertain, which makes sense since both the velocity and the well get larger
for higher states.\\

\section{9.13: }

\section{9.14: Unknown QHO wavefunction}
We are given the wave function $\psi(x,0) = A\left[1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
  2\frac{m\omega}{\hbar}x^2\right]$. Because its highest-order polynomial term is $x^2$, we know
that it must be made up of the $n = 0, 1, 2$ energy states. We now figure out the coefficients for
each:

\begin{align*}
  c_0 &= \int_{-\infty}^{\infty} \phi_0^*\psi dx\\
  &= A\int_{-\infty}^{\infty} \left(\frac{m\omega}{\pi\hbar}\right)^{1/2}e^{-m\omega x^2/\hbar}
  \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
  2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/(2\hbar)}dx\\
  &= 2\\
  c_1 &= A\int_{-\infty}^{\infty} \frac{2\sqrt{\frac{m\omega}{\hbar}x}}{\sqrt{2}}
    \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
    2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
 &= -\frac{3}{\sqrt{2}}\\
 c_2 &= A\int_{-\infty}^{\infty} \frac{\frac{4m\omega}{\hbar}x^2 - 2}{\sqrt{8}}
    \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
    2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
 &= \sqrt{2}\\
\end{align*}

From this we can see that $\psi(x,0) = \sqrt{\frac{21}{2}}
\left(2\ket{0} + -\frac{3}{\sqrt{2}}\ket{1} + \sqrt{2}\ket{2}\right)$. It is then easy to compute
the expectation value: \\

\begin{align*}
  <E> &= \sigma \hbar\omega\left(n+\frac{1}{2}\right)|c_n|^2\\
  &= \frac{2 + \frac{27}{4} + 5}{\frac{21}{2}} = 1.3 eV\\
\end{align*}
\end{document}
